---
title: "hw3"
author: "Shi Tiancheng"
date: "2024-10-16"
output: github_document
---

Problem 2

```{r}
library(tidyverse)
library(tidyr)
library(readr)
library(ggridges)
library(ggplot2)
```
import data and clean
```{r}
covar = read.csv("nhanes_covar.csv")
accel = read.csv("nhanes_accel.csv")
names(covar) = covar[4,]
covar = covar[-c(1:4),]
```

```{r}
covar$SEQN = as.numeric(covar$SEQN)
```

```{r}
nhanes = left_join(covar, accel, by = "SEQN" )
```

```{r}
nhanes_cleaned = filter(nhanes, age >= 21)
nhanes_cleaned = drop_na(nhanes_cleaned)
nhanes_cleaned = mutate(nhanes_cleaned,sex = factor(sex, levels = c(1, 2), labels = c("Male", "Female")),
                        education = factor(education, levels = c(1, 2, 3), 
                        labels = c("Less than high school", 
                                  "High school equivalent", 
                                  "More than high school")))
```

reader-friendly table
```{r}
nhanes_cleaned |>
  group_by(sex, education) |>
  summarize(num = n()) |>
  pivot_wider(names_from = sex, values_from = num) |> 
  knitr::kable()
```

We can find that the high education, the more people. 

```{r}
total_activity = mutate(nhanes_cleaned, total_activity = rowSums(nhanes_cleaned[,c(6:1445)],na.rm=TRUE))
```

plot total activities (y-axis) against age (x-axis)
```{r}
total_activity$age = as.numeric(total_activity$age)
total_activity$total_activity = as.numeric(total_activity$total_activity)
ggplot(total_activity,aes(x=age,y=total_activity)) + 
  geom_point(aes(color = sex)) +
  geom_smooth(se = FALSE) +
  facet_grid(.~ education) +
  labs(
    x = "Age",
    y = "Total Activity"
  ) 
```

Problem 3

import the data
```{r}
Jan_2020 = read.csv("citibike/Jan 2020 Citi.csv") |> mutate(month = "Jan", year = 2020)
Jan_2024 = read.csv("citibike/Jan 2024 Citi.csv") |> mutate(month = "Jan", year = 2024)
July_2020 = read.csv("citibike/July 2020 Citi.csv") |> mutate(month = "July", year = 2020)
July_2024 = read.csv("citibike/July 2024 Citi.csv") |> mutate(month = "July", year = 2024)
```

```{r}
joined_table = bind_rows(Jan_2020, July_2020, Jan_2024, July_2024)
```


reader-friendly table
```{r}
joined_table |>
  group_by(month, year, member_casual) |>
  summarize(num_rides = n()) |>
  pivot_wider(names_from = member_casual, values_from = num_rides) |> 
  knitr::kable()
```
Members are more likely to use sharing bike than casual users.

5 most popular starting stations
```{r}
July_2024 |>
  group_by(start_station_name) |>
  summarize(num_rides = n()) |>
  arrange(desc(num_rides)) |>
  slice_head(n = 5) |>
  knitr::kable()
```

Plot of the effects of day of the week, month, and year on median ride duration.
```{r}
ggplot(joined_table)+geom_boxplot(aes(x=weekdays,y=duration, color = interaction(year, month)))
```
we can find that July 2020 has the largest median among all. Weekdays does not affect a lot, but the median in weekends is larger than weekdays. 

2024 data analysis
```{r}
joined_table_2024 = bind_rows(Jan_2024, July_2024)
```

```{r}
ggplot(joined_table_2024, aes(x = duration, fill = rideable_type)) + 
  geom_histogram() +
  facet_grid(month ~ member_casual)
```
We can find that member ride much more than casual users. Classic bikes are more likely be chosen and July has more rider.
